Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/abstract.tex
some intro nits

\textit{Nail} is an interface generator that greatly reduces the
programmer effort for safely parsing and generating protocols defined
by a Parser-Expression based grammar.  Nail introduces several key
ideas to achieve its goal.  First, Nail uses the protocol grammar to
define not just the data format, but also the internal representation
of the data.  Second, Nail eliminates the notion of semantic actions,
used by existing parser generators, which reduces the expressive power
but allows Nail to both {\em parse} data formats and {\em generate}
them from internal representations, by maintaining a {\em semantic
bijection} between data formats and internal representation.  Third,
Nail introduces {\em dependency fields} and {\em stream transforms} to
capture protocol features such as size and offset fields, checksums,
and compressed data, which are impractical to express in existing
protocol languages.  Using Nail, we implement a complete DNS server in
C in under 300 lines of code and grammar, and a complete {\tt unzip}
program in C in 220 lines of code and grammar, demonstrating that Nail
makes it easy to parse complex real-world data formats.

>>>>

\textit{Nail} is an interface generator that greatly reduces the
programmer effort for safely parsing and generating protocols defined
by a Parser-Expression based grammar.  Nail introduces several key
ideas to achieve its goal.  First, Nail uses the protocol grammar to
define not just the data format, but also the internal object model
of the data.  Second, Nail eliminates the notion of semantic actions,
used by existing parser generators, which reduces the expressive power
but allows Nail to both {\em parse} data formats and {\em generate}
them from the internal object model, by maintaining a {\em semantic
bijection} between the data format and the object model.  Third,
Nail introduces {\em dependency fields} and {\em stream transforms} to
capture protocol features such as size and offset fields, checksums,
and compressed data, which are impractical to express in existing
protocol languages.  Using Nail, we implement a complete DNS server in
C in under 300 lines of code and grammar, and a complete {\tt unzip}
program in C in 220 lines of code and grammar, demonstrating that Nail
makes it easy to parse complex real-world data formats.

### 2014-05-01 paper/osdi14/intro.tex
some intro nits

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so either the programmer has to write potentially unsafe code to deal with such features
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length fields. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations on the external format, such as
compressed data, cannot be implemented at all.

@@

First, Nail grammars define both a format's external representation and an internal object model.
This removes the semantic actions and type declarations that programmers have to write with existing
parser generator. While this somewhat reduces the flexibility of the internal model, it forces the programmer
to clearly separate syntactic validation and semantic processing.

Second, this well-defined internal representation allows Nail to
maintain a {\em semantic bijection} between data formats and their
internal representation.  As a result, this enables Nail to not just
parse input but also generate output from the internal representation,
without requiring the programmer to write additional code.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for Zip files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to handle external data formats.

>>>>

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so the programmer has to either write potentially unsafe code to deal with such features,
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length field. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations, such as
handling compressed data, cannot be implemented at all.

@@

First, Nail grammars define both a format's external representation and an internal object model.
This removes the semantic actions and type declarations that programmers have to write with existing
parser generators. While this somewhat reduces the flexibility of the internal model, it forces the programmer
to clearly separate syntactic validation and semantic processing.

Second, this well-defined internal representation allows Nail to
maintain a {\em semantic bijection} between data formats and their
internal object model.  As a result, this enables Nail to not just
parse input but also generate output from the internal representation,
without requiring the programmer to write additional code.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for Zip files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.
